### 2019-05-13 analysis/paper/cover_letter.Rmd
Add cover letter

---
output: word_document
---

`r format(Sys.time(), '%Y-%m-%d')`

Professor Paul J B Hart  
Fish and Fisheries  
Department of Biology  
University of Leicester  
University Road  
Leicester  
LE1 7RH, UK

Dr. Hart,

We thank you for taking the time to examine our initial idea for our paper "Improving the communication and accessibility of stock assessment using interactive visualization tools". Based on your feedback, we have assembled a paper we feel fits the theme of Ghoti papers, and points towards ways of improving the stock assessment process. We re-iterate that this paper is not technical in nature, but mostly to encourage readers to pursue investigating these interactive visualization tools, and to convince readers that these tools are very much accessible to practicing biologists.

I look forward to your decision.

Sincerely,

Paul Regular  
Fisheries and Oceans Canada  
Northwest Atlantic Fisheries Center  
80 East White Hills, St. John’s, NL  
A1C 5X1, Canada  
E-mail: Paul.Regular@dfo-mpo.gc.ca  
Phone: (709) 772-2067