### 2019-10-16 analysis/dashboards/skeleton.Rmd

---
title: "flexdashboard + plotly + crosstalk"
output: flexdashboard::flex_dashboard
---

```{r setup, include=FALSE}
library(plotly)
library(crosstalk)
lw <- read.csv("length-weight_data.csv")
shared_lw <- SharedData$new(lw, ~record)
```

Column {.tabset}
-------------------------------------

### Length-Weight relationship

```{r}
plot_ly(data = shared_lw) %>% 
    add_markers(x = ~length, y = ~weight, name = "Observed") %>% 
    add_lines(x = ~length, y = ~exp(fit), name = "Predicted")
```

### Residuals vs. fitted values

```{r}
plot_ly(data = shared_lw) %>% 
    add_markers(x = ~fit, y = ~res)
```