### 2019-10-18 analysis/dashboards/getting_started.Rmd
Draft a flexdashboard basics page to the getting_started dashboard

---
title: "Dashboards for stock assessments: Getting started"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: embed
---

<style>
h5 {display: inline;}
</style>

```{r setup, include=FALSE}
library(plotly)
library(crosstalk)
```

Background
=======================================================================

Row {data-height=300}
-----------------------------------------------------------------------

### Dashboards

- Browser-based dashboards and interactive visualizations are becoming increasingly common and accessible  
- With a relatively shallow learning-curve, an [**R**](https://www.r-project.org/) user can use the [**flexdashboard**](https://rmarkdown.rstudio.com/flexdashboard/), [**plotly**](https://plotly-r.com) and [**crosstalk**](https://rstudio.github.io/crosstalk/) packages to generate interactive dashboards for exploring data and models
- The concept is demonstrated on the [Skeleton](#skeleton) page using a simple data-set and linear regression; this can be extended to more complex cases

> Note: this document is meant to be a high-level overview of constructing a dashboard using [**flexdashboard**](https://rmarkdown.rstudio.com/flexdashboard/), [**plotly**](https://plotly-r.com) and [**crosstalk**](https://rstudio.github.io/crosstalk/). This is not a definitive guide to each package nor is this the only approach to constructing a dashboard using R. Please refer to the links provided for more details on how to use the packages highlighted here.

Row {data-height=500}
-----------------------------------------------------------------------

### flexdashboard

- Uses [**rmarkdown**](https://rmarkdown.rstudio.com/) to render a group of related figures, tables and text into a dashboard  
- Layout is flexible and the components automatically re-size to fill the browser and adapt to mobile displays
- Supports a wide range of components, including base plot, ggplots, gauges, tables and [htmlwidgets](http://www.htmlwidgets.org/index.html) such as [**plotly**](https://plotly-r.com), [**leaflet**](https://rstudio.github.io/leaflet/) and [**DT**](https://rstudio.github.io/DT/)
- Optionally use [**shiny**](http://shiny.rstudio.com/) or [**crosstalk**](https://rstudio.github.io/crosstalk/) to bolster interactivity

> Source: https://rmarkdown.rstudio.com/flexdashboard/

### plotly

- A graphing package that works like other R plots except it produces interactive visualizations
- The package allows the user to create interactive web graphics from [**ggplot2**](https://ggplot2.tidyverse.org/) graphs
- Also provides a more 'direct' link to the core [plotly.js](https://plot.ly/javascript/) JavaScript library using syntax inspired by the grammar of graphics

> Source: https://plotly-r.com

### crosstalk

- Enables cross-widget interactions by linking brushing and/or filtering across multiple views
- i.e. Interactions with one plot can affect change in another plot  
- Supports a wide range of [htmlwidgets](http://www.htmlwidgets.org/index.html), such as [**plotly**](https://plotly-r.com), [**leaflet**](https://rstudio.github.io/leaflet/) and [**DT**](https://rstudio.github.io/DT/)

> Source: https://rstudio.github.io/crosstalk/

Skeleton {data-orientation=columns}
=======================================================================

Column {.tabset}
-------------------------------------

### Code

```{r echo = FALSE, comment = ""}
cat(htmltools::includeText("skeleton.Rmd"))
```

> This will generate a simple flexdashboard with two interactive figures (duplicated to the right)

### Details

This skeleton is a verbatim copy of an R markdown file (extension `.Rmd`) set-up to produce a simple `flexdashboard` with interactive `plotly` visuals connected by `crosstalk`. Like most R markdown files, it includes three types of content:

1. A YAML header surrounded by `----`
2. Text following `markdown` formatting
3. R code chunks surrounded by `` ``` ``

The YAML header includes the metadata for the file, such as the document title and output format:

````
---
title: "flexdashboard + plotly + crosstalk"
output: flexdashboard::flex_dashboard
---
````

While only a title and format was specified in the skeleton, many other options are available (e.g. author, date).

The next section is a chunk of R code:

````
```{r setup, include=FALSE} `r ''`
library(plotly)
library(crosstalk)
lw <- read.csv("length-weight_data.csv")
shared_lw <- SharedData$new(lw, ~record)
```
````

This is where the `plotly` and `crosstalk` packages are loaded along with some sample data. The sample data includes length and weight data (columns `length` and `weight`, respectively) along with fitted values and residuals from a length-weight regression (columns called `fit` and `res`, respectively). The `SharedData` function from `crosstalk` package is also used here to generate a data object that can be "shared" across independent plots.

Following this chunk of R code is a level 2 markdown header that tells flexdashboard to introduce a column break and place the subsequent components into separate tabs:

````
Column {.tabset}
-------------------------------------
````

Following this break, independent tabs are defined using three hashtags followed by an optional tab name (level 3 markdown header). This header can be followed by either markdown text or R code. Here, two tabs are generated with `plotly` plots using shared data from `crosstalk`:

````
### Length-Weight relationship

```{r} `r ''`
plot_ly(data = shared_lw) %>% 
    add_markers(x = ~length, y = ~weight, name = "Observed") %>% 
    add_lines(x = ~length, y = ~exp(fit), name = "Predicted")
```

### Residuals vs. fitted values

```{r} `r ''`
plot_ly(data = shared_lw) %>% 
    add_markers(x = ~fit, y = ~res)
```

````

The native syntax of `plotly` was inspired by the grammar of graphics and, as such, its general structure will be familiar to those who have used `ggplot2`. The package has also been structured to be pipe (`%>%`) and `dplyr` friendly, making the code more intuitive and efficient. Though the script ends with the residual vs. fitted plot, the dashboard can easily be extended to include other diagnostic plots, such as a histogram of the residuals.

Once this script is "Knit" (`Ctrl+Shift+K` in Rstudio), a stand-alone html document will be produced with the plots rendered into independent tabs (shown to the right). Clicking a specific point in one plot will highlight the corresponding point in the other plot. In short, `flexdashboard` sets up the structure of the document, `plotly` produces the interactive figures and `crosstalk` connects the plots held in independent tabs.

Of course, this is only a rudimentary overview of what is possible with `flexdashboard`, `plotly` and `crosstalk`. A wide range of layout options are possible using `flexdashboard`, `plotly` can produce more than just scatter and line plots and `crosstalk` can connect various widgets. See the links provided on the [Background](#background) page for more details on each package. The hope here is that this skeleton serves as a starting point from which to build more elaborate dashboards tailored to specific needs.

Column {.tabset}
-------------------------------------

```{r}
lw <- read.csv("length-weight_data.csv")
shared_lw <- SharedData$new(lw, ~record)
```

### Length-Weight relationship

```{r}
plot_ly(data = shared_lw) %>% 
    add_markers(x = ~length, y = ~weight, name = "Observed") %>% 
    add_lines(x = ~length, y = ~exp(fit), name = "Predicted")
```

### Residuals vs. fitted values

```{r}
plot_ly(data = shared_lw) %>% 
    add_markers(x = ~fit, y = ~res)
```

rmarkdown {data-navmenu="Basics"}
=======================================================================

Row {data-height=250}
-------------------------------------

### R markdown

The package `rmarkdown` enables the writing of `markdown` text and `R` code in the same document (extension `.Rmd`). Most R markdown files includes three types of content:

1. A YAML header surrounded by `----`
2. Text following `markdown` formatting
3. R code chunks surrounded by `` ``` ``

Row {data-height=600}
-------------------------------------

### YAML

The first step in setting up an R markdown file is to define a YAML header. In the [skeleton](#skeleton), it looks like this:

````
---
title: "flexdashboard + plotly + crosstalk"
output: flexdashboard::flex_dashboard
---
````

This is where the output type is defined along with other items such as title, author, date, etc. A `flex_dashboard` is requested here. There are are growing number of output types supported by R markdown (e.g. html_document, pdf_document, word_document).

### Markdown

Next, write plain text using markdown syntax to describe how to format the text in the final document.

- `Plain text` simply translates to Plain text in the default font of the output document.
- Italics and bold are specified using asterisk and/or underscores:
    - `*italic*` or `_italic_` $\rightarrow$ *italic*
    - `**bold**` or `__bold__` $\rightarrow$ **bold**
    - `**_bold-italic_** ` $\rightarrow$ **_bold-italic_** 
- Equations can be specified using LaTeX syntax:
    - `$W = \alpha L ^{\beta}$` $\rightarrow$ $W = \alpha L ^{\beta}$
- Ordered and unordered lists are generated using numbers and dashes or asterisk, respectively:
    - `1. ordered item 1` $\rightarrow$ 1. &nbsp; ordered item 1
    - `- unordered item` $\rightarrow$ • &nbsp; unordered item
- Headers are define using hashtags:
   - `##### Header 5` $\rightarrow$ <h5> Header 5 </h5>
   - Note that markdown headers with three hashtags or less have specific formatting functions in `flexdashboard`

### R

Finally, mix in `R` code by surrounding chunks of code using one backtick for inline code (e.g.`` `r knitr::inline_expr("1+1")` `` will print 2 in the output) or three backticks to run several lines of code and/or display a table or plot. The chunk below will print the top three rows of the `cars` data-set:

````
```{r} `r ''`
head_cars <- head(cars, 3)
head_cars
```
````
Resulting in the following output:

```{r}
head_cars <- head(cars, 3)
head_cars
```

flexdashboard {data-navmenu="Basics"}
=======================================================================

Row {data-height=200}
-------------------------------------

### flexdashboard

The `flexdashboard` package can be used to render groups of related text, figures and tables into a dashboard. Using a combination of `markdown` syntax and `R` code (i.e. R markdown), this package facilitates a wide range of layout options and each component can include output from packages such as `plotly`, `leaflet`, `ggplot2`, and so on. The package also integrates nicely with `shiny` and `crosstalk`, providing options for increasing the interactivity of the dashboard. Some layout and component options are highlited below.

Row {data-height=600}
-------------------------------------

### Layout

Individual charts are defined using markdown's level 3 header (`### Chart title`) and they are, by default, stacked vertically within columns defined using markdown's level 2 header (`---------`). The code below will create a `flexdashboard` with two stacked columns and three charts:

````
---
title: "Layout example 1"
output: flexdashboard::flex_dashboard
---

Column
-------------------------------------

### Chart 1

Column
-------------------------------------

### Chart 2

### Chart 3

````

The output from this example has been duplicated [here](#layout-example-1).

Alternatively, charts can be organized by row by modifying the YAML header:

````
---
title: "Layout example 2"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
---

Row
-------------------------------------

### Chart 1

Row
-------------------------------------

### Chart 2

### Chart 3

````
The output from this example has been duplicated [here](#layout-example-2).

### Components

The components of a dashboard can hold a wide range of outputs. The code below generates a mixture of tabular and graphical output. The table is generated following `markdown` syntax with inline `R` code and the interactive plot is generated using `plotly`.

````
---
title: "Components example"
output: flexdashboard::flex_dashboard
---

Column {data-width=300}
-------------------------------------

### Table

Summary statistics of the `volcano` data-set

| Statistic     | Elevation                   |
|:------------- |:--------------------------- |
| Min           | `r knitr::inline_expr('min(volcano)')`            | 
| Median        | `r knitr::inline_expr('median(volcano)')`         |
| Mean          | `r knitr::inline_expr('round(mean(volcano))')`    |
| Max           | `r knitr::inline_expr('max(volcano)')`            |

Column {data-width=700}
-------------------------------------

### Plot

```{r} `r ''`
plotly::plot_ly(z = volcano, type = "surface")
```

````

The output from this example has been duplicated [here](#components-example). Note the use of the `data-width` attribute to make the table chart relatively narrow.

Layout example 1 {data-orientation=columns .hidden}
=======================================================================

Column
-------------------------------------

### Chart 1

Column
-------------------------------------

### Chart 2

### Chart 3

Layout example 2 {data-orientation=rows .hidden}
=======================================================================

Row
-------------------------------------

### Chart 1

Row
-------------------------------------

### Chart 2

### Chart 3

Components example {data-orientation=columns .hidden}
=======================================================================

Column {data-width=300}
-------------------------------------

### Table

Summary statistics of the `volcano` data-set

| Statistic     | Elevation                   |
|:------------- |:--------------------------- |
| Min           | `r min(volcano)`            |
| Median        | `r median(volcano)`         |
| Mean          | `r round(mean(volcano))`    |
| Max           | `r max(volcano)`            |

Column {data-width=700}
-------------------------------------

### Plot

```{r}
plotly::plot_ly(z = volcano, type = "surface")
```