### 2020-03-14 inst/rmd/fishery.Rmd

Landings {data-navmenu="Fishery"}
================================================================================

Row {.tabset .tabset-dropdown}
--------------------------------------------------------------------------------

### Landings

```{r}

plot_landings(fit, cols = c(pred_col, obs_col))

```

> Y-axes in log scale

### Numbers at age

```{r}

d <- tidy_mats(list(obs = t(inputs$catch.num[as.character(years), as.character(ages)]),
                    pred = fit$rep$pred_catch),
               ages = ages, years = years)
d$cohort <- factor(d$year - d$age)
d$age <- factor(d$age)

facet_obs_pred(d, by = "age", type = "log",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Proportion",
               obs_col = obs_col, pred_col = pred_col) %>%
  layout(yaxis = list(rangemode = "tozero"))

```

> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration

### Numbers at age (frame)

```{r}
plot_obs_pred(data = d, showlegend = TRUE, type = "log",
              frame = "age", ylab = "Numbers",
              obs_col = obs_col, pred_col = pred_col)

```

> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration

Row {.tabset}
--------------------------------------------------------------------------------

### Proportions at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$catch_prop,
                    pred = fit$rep$pred_prop_catch),
               ages = ages, years = years)
d$cohort <- factor(d$year - d$age)
d$age <- factor(d$age)

facet_obs_pred(d, by = "age",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Proportion",
               obs_col = obs_col, pred_col = pred_col) %>%
  layout(yaxis = list(rangemode = "tozero"))

```

### Proportions at age (frame)

```{r}

plot_obs_pred(data = d, showlegend = TRUE, type = "linear",
              frame = "age", ylab = "Proportion",
              obs_col = obs_col, pred_col = pred_col)

```

### Residuals

```{r}

d <- tidy_mats(list(res = fit$rep$CA_resid,
                    fit = fit$rep$pred_crl),
               ages = ages[-length(ages)], years = years)
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = obs_col, line_col = pred_col)

```

> based on continuation ratio logits

Selectivity {data-navmenu="Fishery"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Proportions at age surface

```{r}
plot_surface(mat = fit$rep$pred_prop_catch, ages = ages, years = years, name = "P")
```

<!-- ### Numbers at age surface -->

<!-- ```{r} -->
<!-- plot_surface(mat = t(inputs$catch.num[as.character(years), as.character(ages)]),  -->
<!--              ages = ages, years = years, name = "N") -->
<!-- ``` -->

<!-- > Note: model fits to composition - this plot is included for illustration -->

Row {.tabset}
--------------------------------------------------------------------------------

### Fishery selectivity

```{r}

S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x))
plot_surface(S, ages = ages, years = years, name = "S")

```