### 2020-03-14 inst/rmd/projections.Rmd

Assumptions  {data-navmenu="Projections" .storyboard}
================================================================================

### Assume the same selectivity curve as estimated by NCAM in the terminal year and apply catch multipliers

```{r}

Cp <- fit$tmb.data$Cp
Cp_string <- text_c(Cp)
Lp <- round(Cp * tail(fit$rep$pred_catch_wt, 1))
Lp_string <- paste(text_c(Lp), "t")

tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "F") %>%
  filter(year >= max(years) - 3) %>%
  plot_ly(x = ~age, y = ~F, color = ~factor(year), colors = rev(cols)) %>%
  add_lines(alpha = 0.5) %>%
  filter(year == max(years)) %>%
  add_lines(showlegend = FALSE, line = list(width = 3)) %>%
  layout(xaxis = list(title = "Age"))

```

***
- The following catch multipliers were applied: `r Cp_string`
- This translates to `r Lp_string` of catch in the projection years

### Assume natural mortality converges on the long-term mean, excluding the M spike between 1991 - 1994

```{r}

if (all(fit$tmb.data$Mp_mult != 1)) stop("TODO: make the mortality multiplier projection documentation dynamic")

Mpe <- fit$rep$Mpe_matrix
Mproj <- replicate(fit$tmb.data$Yp, fit$tmb.data$pnm)
bpe <- fit$rep$bpe[, (fit$tmb.data$Y + 1):(fit$tmb.data$Y + fit$tmb.data$Yp)]
Mproj <- Mproj * exp(bpe)
Mproj <- cbind(Mpe, Mproj)
proj_years <- fit$tmb.data$proj_years

plot_surface(mat = Mproj, ages = ages, years = c(years, proj_years), name = "M")

```

### Assume mean recruitment from the last three years

```{r}

last_three <- tail(fit$tmb.data$years, 3)
mean_rec <- exp(mean(tabs$log_recruitment$est[tabs$log_recruitment$year %in% last_three]))
mean_rec <- data.frame(year = c(last_three, last_three[3] + fit$tmb.data$Yp), 
                       mean_rec = mean_rec * 1000, grp = "mean")
trans_est(tabs$log_recruitment, scale = 0.001) %>%
  plot_trend(ylab = "Recruitment", col = pred_col) %>% 
  add_lines(data = mean_rec, x = ~year, y = ~mean_rec, name = "mean",
            color = I("black"), line = list(width = 1)) %>% 
  add_text(x = max(years) + 1.2, y = mean_rec$mean_rec[1], 
           text = "mean recruitment", 
           textposition = "bottom right",
           color = I("black")) %>%
  layout(xaxis = list(range = c(min(years), max(years) + fit$tmb.data$Yp)))

```

### Assume averages weights and proportions mature from the last three years

```{r}

mat_p <- tidy_mat(fit$tmb.data$mat, ages = factor(ages), years = years, value = "mat") %>%
  filter(year >= max(years) - 2) %>%
  plot_ly(x = ~age, y = ~mat, color = ~factor(year), colors = rev(cols)) %>%
  add_lines(alpha = 0.5) %>%
  group_by(age) %>%
  summarise(mat = mean(mat), year = "3y mean") %>%
  add_lines(line = list(width = 3)) %>%
  layout(xaxis = list(title = "Age"),
         yaxis = list(title = "Proportion mature"))

wt_p <- tidy_mat(fit$tmb.data$weight, ages = factor(ages), years = years, value = "weight") %>%
  filter(year >= max(years) - 2) %>%
  plot_ly(x = ~age, y = ~weight, color = ~factor(year), colors = rev(cols),
          showlegend = FALSE) %>%
  add_lines(alpha = 0.5) %>%
  group_by(age) %>%
  summarise(weight = mean(weight), year = "3y mean") %>%
  add_lines(line = list(width = 3)) %>%
  layout(xaxis = list(title = "Age"),
         yaxis = list(title = "Weight (kg)"))

subplot(mat_p, wt_p, titleY = TRUE, titleX = TRUE)

```

Past projections {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`}
================================================================================

### Catch multiplier projections for SSB / Blim from the last RAP

```{r, eval = !is.null(tmp$last)}

d <- trans_est(last_tabs$Cproj_log_ssb_lrp)
d$scenario <- factor(d$scenario)
d$type <- "Projected"

past_d <- trans_est(last_tabs$log_ssb_lrp)
past_d <- past_d[past_d$year != max(past_d$year), ] # drop the last year here (it's in the projection output)
past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s))
past_d <- do.call(rbind, past_d)
past_d$type <- "Estimated"
d <- rbind(past_d, d)
xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1)
ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])

current_d <- trans_est(tabs$log_ssb_lrp)
current_d <- lapply(levels(d$scenario), function(s) data.frame(current_d, scenario = s))
current_d <- do.call(rbind, current_d)
current_d$model <- "current RAP"
current_d$type <- "Estimated"
d$model <- "last RAP"
d <- rbind(d, current_d)

plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", group = "model",
          vline = max(last$tmb.data$years) + 1, hline = 1, col = cols, showlegend = TRUE,
          darken_proj = TRUE)

```

Retro projections {data-navmenu="Projections" `r if (is.null(fit$retro)) ".hidden"`}
================================================================================

### Catch multiplier projections for SSB / Blim from retrospective fits

```{r, eval = !is.null(fit$retro)}

d <- trans_est(retro_tabs$Cproj_log_ssb_lrp)
d$terminal_year <- as.numeric(as.character(d$terminal_year))
d$scenario <- factor(d$scenario)
d$type <- "Projected"

past_d <- trans_est(retro_tabs$log_ssb_lrp)
past_d$terminal_year <- as.numeric(as.character(past_d$terminal_year))
past_d <- past_d[past_d$year <= past_d$terminal_year, ] # drop the last year (it's in the projection output)
past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s))
past_d <- do.call(rbind, past_d)
past_d$type <- "Estimated"
d <- rbind(past_d, d)
d <- d[order(d$scenario, d$terminal_year, d$year), ]

# xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1)
# ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])

plot_proj(data = d, ylab = "SSB/Blim", group = "terminal_year",
          hline = 1, col = rev(viridis::viridis(length(unique(d$terminal_year)))), 
          showlegend = TRUE, markers = FALSE, darken_proj = TRUE)

```

Results {data-navmenu="Projections"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Catch multiplier projections for SSB~yp~ / SSB~`r max(years) + 1`~

```{r}

d <- trans_est(tabs$Cproj_log_ssb_ratio)
d[, c("est", "lwr", "upr")] <- (d[, c("est", "lwr", "upr")] - 1) * 100
d$scenario <- factor(d$scenario)
xlim <- c(max(years) + 0.9, max(years) + 4.1)
ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])

plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "Percent growth", ticksuffix = "%",
          vline = max(years) + 1, col = pred_col) %>%
  layout(yaxis = list(tickvals = seq(-200, 200, by = 25)))

```

### SSB / Blim

```{r}

d <- trans_est(tabs$Cproj_log_ssb_lrp)
d <- d[d$year != max(years_plus1), ] # drop the first year (it's in the time series)
d$type <- "Projected"
d$scenario <- factor(d$scenario)

## Append full time series to projections
past_d <- trans_est(tabs$log_ssb_lrp)
past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s))
past_d <- do.call(rbind, past_d)
past_d$type <- ifelse(past_d$year == max(years_plus1), "Projected", "Estimated")
d <- rbind(past_d, d)
xlim <- c(max(years) - 2.1, max(years) + 4.1)
ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])

plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim",
          vline = max(years) + 1, hline = 1, col = pred_col, darken_proj = TRUE)

```

Row {.tabset}
--------------------------------------------------------------------------------

### Prob(SSB~yp~ &ge; SSB~`r max(years) + 1`~)

```{r}

growth <- tabs$Cproj_log_ssb_ratio
growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd)

d <- growth[growth$year != max(years) + 1, ]
d$scenario <- factor(d$scenario)
plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>%
  add_heatmap() %>%
  add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2),
                  showarrow = FALSE, font = list(color = "white")) %>%
  colorbar(title = "P", limits = c(0, 1)) %>%
  layout(yaxis = list(title = "Catch multiplier"),
         xaxis = list(title = "Year", 
                      tickvals = min(d$year):max(d$year)))

```

### Prob(SSB~yp~ &ge; B~lim~)

```{r}

growth <- tabs$Cproj_log_ssb_lrp
growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd)

d <- growth[growth$year != max(years) + 1, ]
d$scenario <- factor(d$scenario)
plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>%
  add_heatmap() %>%
  add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2),
                  showarrow = FALSE, font = list(color = "white")) %>%
  colorbar(title = "P", limits = c(0, 1)) %>%
  layout(yaxis = list(title = "Catch multiplier"),
         xaxis = list(title = "Year",
                      tickvals = min(d$year):max(d$year)))

```