### 2020-03-14 inst/rmd/surveys.Rmd

RV Survey  {data-navmenu="Surveys"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### RV survey index

```{r}

obs <- colSums(fit$tmb.data$est_wt * fit$tmb.data$RV_index * fit$tmb.data$trawl_units)
pred <- colSums(fit$tmb.data$est_wt * exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units)
obs[obs == 0] <- NA
pred[pred == 0] <- NA
d <- data.frame(year = years, obs = obs, pred = pred)

# ## RV biomass
# obs <- colSums(fit$tmb.data$RV_index * fit$tmb.data$trawl_units * fit$tmb.data$weight)
# pred <- colSums(exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units * fit$tmb.data$weight)
# d <- data.frame(year = years, obs = obs, pred = pred)
# write.table(d, file = "clipboard", sep = "\t", row.names = FALSE, quote = FALSE)

plot_obs_pred(d, obs_col = obs_col, pred_col = pred_col, ylab = "RV survey index",
              showlegend = TRUE, type = "log")

```

> Y-axes in log scale

### Index at age

```{r}

d <- frame_d <- tidy_mats(list(obs = fit$tmb.data$RV_index,
                               pred = exp(fit$rep$Elog_RV_index),
                               wt = fit$tmb.data$est_wt),
                          ages = ages, years = years)
d$obs[d$wt == 0] <- NA
d$pred[d$wt == 0] <- NA
d$cohort <- frame_d$cohort <- factor(d$year - d$age)
d$age <- frame_d$age <- factor(d$age)

facet_obs_pred(d, by = "age",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Mean numbers per tow",
               obs_col = obs_col, pred_col = pred_col, type = "log")

```

> Y-axes in log scale

### Index at age (frame)

```{r}
plot_obs_pred(data = frame_d, showlegend = TRUE, 
              type = "log", frame = "age",
              obs_col = obs_col, pred_col = pred_col,
              ylab = "Mean numbers per tow")
```

> Y-axes in log scale

Row {.tabset}
--------------------------------------------------------------------------------

### Residuals

```{r}

d <- tidy_mats(list(res = fit$rep$RV_resid,
                    keep = fit$tmb.data$ind_zero == 0 & fit$tmb.data$est_wt == 1,
                    fit = fit$rep$Elog_RV_index),
               ages = ages, years = years)
d <- d[d$keep, ]
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = obs_col, line_col = pred_col)

```

### Residuals (bubble)

```{r}

d <- tidy_mats(list(res = fit$rep$RV_resid,
                    zero = fit$tmb.data$ind_zero,
                    wt = fit$tmb.data$est_wt),
               ages = ages, years = years)
d <- d[d$wt == 1, ]
bubble_plot(d, pos_col = pos_col, neg_col = neg_col, na_col = obs_col)

```

SN Survey {data-navmenu="Surveys"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Sentinel gillnet survey index

```{r}

obs <- colSums(fit$tmb.data$SN_GN_index)
pred <- colSums(exp(fit$rep$Elog_SN_GN_index))
d <- data.frame(year = fit$tmb.data$SN_years, obs = obs, pred = pred)

plot_obs_pred(d, obs_col = obs_col, pred_col = pred_col, ylab = "Sentinel survey index",
              showlegend = TRUE, type = "log")

```

> Y-axes in log scale

### Index at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$SN_GN_index,
                    pred = exp(fit$rep$Elog_SN_GN_index)),
               ages = fit$tmb.data$SN_ages, years = fit$tmb.data$SN_years)
d$cohort <- factor(d$year - d$age)
d$age <- factor(d$age)

facet_obs_pred(d, by = "age",
               which = c(3, 7, 4, 8, 5, 9, 6, 10), # done to order by column
               nrows = 4, ylab = "Sentinel survey index",
               obs_col = obs_col, pred_col = pred_col, type = "log")

```

> Y-axes in log scale

### Index at age (frame)

```{r}
plot_obs_pred(data = d, showlegend = TRUE, 
              type = "log", frame = "age",
              obs_col = obs_col, pred_col = pred_col,
              ylab = "Sentinel survey index")
```

> Y-axes in log scale

Row {.tabset}
--------------------------------------------------------------------------------

### Residuals

```{r}

d <- tidy_mats(list(res = fit$rep$log_SN_GN_resid,
                    fit = fit$rep$Elog_SN_GN_index),
               ages = fit$tmb.data$SN_ages,
               years = fit$tmb.data$SN_years)
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = obs_col, line_col = pred_col)

```

### Residuals (bubble)

```{r}

d <- tidy_mat(fit$rep$log_SN_GN_resid,
              ages = fit$tmb.data$SN_ages,
              years = fit$tmb.data$SN_years, value = "res")
bubble_plot(d, pos_col = pos_col, neg_col = neg_col)

```

SS Survey {data-navmenu="Surveys"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Smith sound acoustic biomass estimates

```{r}

d <- data.frame(year = fit$tmb.data$SS_bms_years,
                obs = fit$tmb.data$SSbms,
                pred = fit$rep$ESSbms)
plot_obs_pred(d, obs_col = obs_col, pred_col = pred_col, ylab = "Biomass (t)",
              showlegend = TRUE)

```

### Numbers at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$SS_agecomps,
                    pred = fit$rep$ESS_agecomps),
               ages = fit$tmb.data$SS_ages,
               years = fit$tmb.data$SS_age_years)

facet_obs_pred(d, by = "age",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Numbers caught", rangemode = "tozero",
               obs_col = obs_col, pred_col = pred_col)

```

### Numbers at age (frame)

```{r}
plot_obs_pred(data = d, showlegend = TRUE, 
              frame = "age",
              obs_col = obs_col, pred_col = pred_col,
              ylab = "Numbers caught")
```

Row {.tabset}
--------------------------------------------------------------------------------

### Residuals

```{r}

d <- tidy_mats(list(res = fit$rep$SS_agecomps_resid_std,
                    fit = fit$rep$ESS_agecomps),
               ages = fit$tmb.data$SS_ages,
               years = fit$tmb.data$SS_age_years)
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = obs_col, line_col = pred_col)

```

### Residuals (bubble)

```{r}

d <- tidy_mat(fit$rep$SS_agecomps_resid_std,
              ages = fit$tmb.data$SS_ages,
              years = fit$tmb.data$SS_age_years, value = "res")
bubble_plot(d, pos_col = pos_col, neg_col = neg_col)

```

Catchability {data-navmenu="Surveys"}
================================================================================

Row
--------------------------------------------------------------------------------

### RV survey catchability

```{r}

q <- exp(fit$rep$log_RV_Qa + log(fit$tmb.data$trawl_units))
plot_surface(q, ages = ages, years = years, name = "q")

```

Row {.tabset}
--------------------------------------------------------------------------------

### Proportion of the stock in Smith Sound

```{r}
p <- matrix(0, nrow = length(ages), ncol = length(years),
            dimnames = list(ages, years))
ss_p <- fit$rep$SSD
ss_years <- fit$tmb.data$SS_years
ss_ages <- fit$tmb.data$SS_ages
rownames(ss_p) <- ss_ages
colnames(ss_p) <- ss_years
p[as.character(ss_ages), as.character(ss_years)] <- ss_p
plot_surface(p, ages = ages, years = years, name = "p")

```

### Sentinel gillnet survey catchability

```{r}

log_qy <- t(replicate(length(fit$tmb.data$SN_ages), fit$rep$SN_GN_Qy))
log_qa <- fit$rep$log_SN_GN_Qa
q <- exp(log_qy + log_qa)
plot_surface(q, ages = fit$tmb.data$SN_ages,
             years = fit$tmb.data$SN_years,
             name = "q")

```