### 2020-03-14 inst/rmd/tables.Rmd

Inputs {data-navmenu="Tables"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Landings

```{r}
kable(inputs$landings, caption = "Reported landings (t) of cod from NAFO Divs. 2J+3KL from 1959 onward")
```

### Catch at age

```{r}
kable(inputs$catch.num, caption = "Catch numbers-at-age (thousands) for the commercial cod fishery in NAFO Divs. 2J+3KL.")
```

<!-- ### Monthly landings -->

<!-- ```{r} -->
<!-- d <- inputs$catch.month -->
<!-- dimnames(d) <- list(year = rownames(d), month = month.name) -->
<!-- kable(d, caption = "Monthly reported landings of cod from NAFO Divs. 2J+3KL.") -->
<!-- ``` -->

### RV index

```{r}
d <- data.table::dcast(inputs$index, year ~ age, value.var = "index")
mat <- d[, as.character(2:14)]
rownames(mat) <- d$year
kable(mat, caption = "Mean number of cod per tow at age in the index strata for the autumn DFO RV bottom-trawl surveys of NAFO Divs. 2J+3KL.", digits = 3)
```

### SN index

```{r}
d <- data.table::dcast(inputs$sentinel, year ~ AGE, value.var = "EST")
mat <- d[, as.character(3:10)]
rownames(mat) <- d$year
kable(mat, caption = 'Standardized age-disaggregated sentinel gill-net (5 1/2" mesh) catch rates (fish / net) of cod in NAFO Divs. 2J+3KL.', digits = 3)
```

### SS index

```{r}
d <- inputs$SS.bms[, c("Year", "month", "Biomass", "se2")]
names(d) <- c("Year", "Month", "Biomass", "SE^2^")
kable(d, caption = "Acoustic biomass estimates (t) for Smith Sound, Trinity Bay, from Rose et al. (2011).")
```

### SS age composition

```{r}
kable(inputs$SS_agecomps, caption = "Age composition of cod sampled during acoustic surveys in Smith Sound during 1995-2003. For details see Rose et al. 2011.")
```

### Weight at age

```{r}
d <- inputs$pwts
d <- d[d$year <= max(inputs$index$year), ]
d <- data.table::dcast(d, year ~ age, value.var = "weight")
mat <- d[, as.character(2:14)]
rownames(mat) <- d$year
kable(mat, caption = "Beginning of year weight-at-age estimates (kg) from a generalized Von Bertalanffy (VonB2) growth model described in Cadigan (2016) fitted by cohort to average weight at age data for cod from autumn bottom-trawl surveys in Division 2J+3KL.", digits = 3)
```

### Maturity at age

```{r}
d <- inputs$mats
d <- d[d$year <= max(inputs$index$year), ]
d <- data.table::dcast(d, year ~ age, value.var = "mat")
mat <- d[, as.character(2:14)]
rownames(mat) <- d$year
kable(mat, caption = "Estimated proportions mature for female cod from NAFO Div. 2J+3KL from DFO RV autumn bottom trawl surveys. Estimates were obtained from a cohort-specific binomial logistic regression model fitted to observed proportions mature at age.", digits = 3)
```

### Tag summary

```{r}
d <- inputs$mtag1 %>% 
  group_by(rel_year, exp) %>% 
  summarise(NAFO = unique(div),
            Nrel = sum(unique(released)), 
            Ncap = sum(recaptured))
names(d) <- c("Year", "Experiment", "NAFO division", "N~released~", "N~recaptured~")
kable(d, caption = "Number of tagged cod released and recaptured by experiment.")
```

Settings {data-navmenu="Tables"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Catch bounds

```{r}
d <- data.frame(Year = fit$tmb.data$years, Lower = fit$tmb.data$LRL, Upper = fit$tmb.data$LRU)
kable(d, caption = "Lower and upper catch multipliers used for generating catch bounds.")
```

### Baseline M

```{r}
m <- fit$tmb.data$M_matrix
rownames(m) <- ages
colnames(m) <- years
kable(t(m), caption = "Baseline levels of natural mortality (M) at age.")
```

Output {data-navmenu="Tables"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Parameter estimates

```{r}

par_tab <- tidy_pars(fit)
rownames(par_tab) <- NULL
options(knitr.kable.NA = '')
kable(par_tab, digits = 3, caption = "Estimates of model fits, population parameters and variance parameters, with coefficients of variation.")

```

> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial

### Parameter estimates (retro)

```{r}

par_tabs <- lapply(names(fit$retro), function(nm) {
  d <- data.frame(tyear = as.numeric(nm), tidy_pars(fit$retro[[nm]]))
  d$id <- seq(nrow(d))
  d
})
par_tabs <- do.call(rbind, par_tabs)
final_par_tab <- data.frame(tyear = max(years), par_tab)
final_par_tab$id <- seq(nrow(final_par_tab))
par_tabs <- rbind(par_tabs, final_par_tab)
par_tabs <- par_tabs[order(par_tabs$id, par_tabs$tyear), ]
par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
par_tabs$Quantity <- as.character(par_tabs$Quantity)
par_tabs$Quantity[duplicated(par_tabs$Quantity)] <- ""
par_tabs$Symbol <- as.character(par_tabs$Symbol)
par_tabs$Symbol[duplicated(par_tabs$Symbol)] <- ""
par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
rownames(par_tabs) <- NULL
names(par_tabs)[names(par_tabs) == "tyear"] <- "Terminal Year"

kable(par_tabs, digits = 3, caption = "Estimates of retrospective model fits, population parameters and variance parameters, with coefficients of variation.")

```

> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial

### Stock size

```{r}

tab_name <- c("log_abundance" = "Abundance (millions)", 
              "log_recruitment" = "Recruits (age 2; millions)", 
              "log_biomass" = "Biomass (kt)", 
              "log_ssa" = "Spawners (millions)", 
              "log_ssb" = "SSB (kt)", 
              "log_ssb_lrp" = "SSB/Blim n(%)")

nms <- names(tab_name)
status_tab <- trans_est(tabs[[nms[1]]], scale = 1000)[, c("year", "est", "lwr", "upr")]
names(status_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))

for (n in nms[-1]) {
  s <- ifelse(n == "log_ssb_lrp", 0.01, 1000)
  tab <- trans_est(tabs[[n]], scale = s)[, c("year", "est", "lwr", "upr")]
  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
  status_tab <- merge(status_tab, tab, by = "Year")
}

knitr::kable(status_tab, digits = 0, caption = "Northern cod stock size estimates with lower (L) and upper (U) 95% confidence intervals (CI)")

```

### Rates

```{r}

tab_name <- c("log_aveZ" = "Average Z (ages 5+)", 
              "log_aveM" = "Average M (ages 5+)", 
              "log_aveF" = "Average F (ages 5+)")

nms <- names(tab_name)
rate_tab <- trans_est(tabs[[nms[1]]])[, c("year", "est", "lwr", "upr")]
names(rate_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))

for (n in nms[-1]) {
  tab <- trans_est(tabs[[n]], scale = 1)[, c("year", "est", "lwr", "upr")]
  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
  rate_tab <- merge(rate_tab, tab, by = "Year")
}

knitr::kable(rate_tab, digits = 3, caption = "Northern cod stock mortality rate estimates with lower (L) and upper (U) 95% confidence intervals (CI)")

```

### Abundance-at-age

```{r}
N_matrix <- rbind(fit$rep$N_matrix, fit$rep$abundance)
dimnames(N_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
knitr::kable(t(N_matrix) / 1000, caption = "Abundance-at-age estimates (millions).", digits = 2)
```

### Mature abundance-at-age

```{r}
ssa_matrix <- rbind(fit$rep$ssa_matrix, fit$rep$ssa)
dimnames(ssa_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
knitr::kable(t(ssa_matrix) / 1000, caption = "Mature abundance-at-age estimates (millions).", digits = 2)
```

### Biomass-at-age

```{r}
B_matrix <- rbind(fit$rep$biomass_matrix, fit$rep$biomass)
dimnames(B_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
knitr::kable(t(B_matrix) / 1000, caption = "Biomass-at-age estimates (kt).", digits = 2)
```

### Mature biomass-at-age

```{r}
ssb_matrix <- rbind(fit$rep$ssb_matrix, fit$rep$ssb)
dimnames(ssb_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
knitr::kable(t(ssb_matrix) / 1000, caption = "Mature biomass-at-age estimates (kt).", digits = 2)
```

<!-- ### Exploitable biomass -->

<!-- ```{r} -->
<!-- S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x)) -->
<!-- B <- fit$rep$N_matrix[, seq(ncol(S))] * fit$tmb.data$midyw -->
<!-- E <- B * S -->
<!-- ebms_matrix <- rbind(E, colSums(E)) -->
<!-- dimnames(ebms_matrix) <- list(age = c(ages, "Total"), year = years) -->
<!-- knitr::kable(t(ebms_matrix) / 1000, -->
<!--              caption = "Exploitable biomass estimates (kt).", digits = 3) -->
<!-- ``` -->

### Predicted catch

```{r}
pred_catch_wt_matrix <- rbind(fit$rep$pred_catch_wt_matrix, fit$rep$pred_catch_wt)
dimnames(pred_catch_wt_matrix) <- list(age = c(ages, "Total"), year = years)
knitr::kable(t(pred_catch_wt_matrix) / 1000, 
             caption = "NCAM predicted catch at age (kt).", digits = 3)
```

### Z-at-age

```{r}
Z_matrix <- fit$rep$Z_matrix
dimnames(Z_matrix) <- list(age = ages, year = years)
knitr::kable(t(Z_matrix), caption = "Z-at-age estimates.", digits = 3)
```

### M-at-age

```{r}
pe_ages <- fit$tmb.data$pe_ages
M_matrix <- fit$rep$Mpe_matrix
dimnames(M_matrix) <- list(age = ages, year = years)
M_matrix <- M_matrix[as.character(pe_ages), ]
rownames(M_matrix) <- c(as.character(head(pe_ages, -1)), paste0(tail(pe_ages, 1), "+"))
knitr::kable(t(M_matrix), caption = "M-at-age estimates.", digits = 3)
```

### F-at-age

```{r}
F_matrix <- fit$rep$F_matrix
dimnames(F_matrix) <- list(age = ages, year = years)
knitr::kable(t(F_matrix), caption = "F-at-age estimates.", digits = 3)
```