### 2020-03-14 inst/rmd/trends.Rmd

Retro {data-navmenu="Trends" `r if (is.null(fit$retro)) ".hidden"`}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Stock status

```{r, eval = !is.null(fit$retro)}

bms_plot <- trans_est(retro_tabs$log_ssb) %>%
  plot_retro(ylab = "SSB (t)", lwd = 1, col = pred_col)

abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
  plot_retro(ylab = "Recruitment (age 2)", lwd = 1, col = pred_col)

blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
  plot_retro(ylab = "SSB/Blim", lwd = 1, col = pred_col) %>%
  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none")

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

### Mohn's rho

```{r, eval = !is.null(fit$retro)}

bms_plot <- trans_est(retro_tabs$log_ssb) %>%
  plot_mohns(col = pred_col, ylab = "SSB")

abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
  plot_mohns(ylab = "Recruitment (age 2)", col = pred_col)

blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
  plot_mohns(ylab = "SSB/Blim", col = pred_col)

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
  layout(annotations = left_label("Relative difference", offset = 50),
         margin = list(l = 75))

```

> Relative difference = (retro estimate - current estimate) / current estimate; Mohn's rho = mean(relative difference)

### Difference

```{r, eval = !is.null(fit$retro)}

bms_plot <- retro_tabs$log_ssb %>%
  plot_retro_deltas(in_col = pred_col, ylab = "log(SSB)")

abn_plot <- retro_tabs$log_recruitment %>%
  plot_retro_deltas(ylab = "log(Recruitment)", in_col = pred_col)

blim_plot <- retro_tabs$log_ssb_lrp %>%
  plot_retro_deltas(ylab = "log(SSB/Blim)", in_col = pred_col)

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
  layout(annotations = left_label("Difference", offset = 50),
         margin = list(l = 75))

```

> Difference = retro estimate - current estimate; approximate 95% intervals based on retro standard deviation + current standard deviation

Row {.tabset}
--------------------------------------------------------------------------------

### Mortality Rates

```{r, eval = !is.null(fit$retro)}

z_plot <- trans_est(retro_tabs$log_aveZ) %>%
  plot_retro(ylab = "Average Z (ages 5+)", lwd = 1, col = pred_col)

m_plot <- trans_est(retro_tabs$log_aveM) %>%
  plot_retro(ylab = "Average M (ages 5+)", lwd = 1, col = pred_col)

f_plot <- trans_est(retro_tabs$log_aveF) %>%
  plot_retro(ylab = "Average F (ages 5+)", lwd = 1, col = pred_col)

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

### Mohn's rho

```{r, eval = !is.null(fit$retro)}

z_plot <- trans_est(retro_tabs$log_aveZ) %>%
  plot_mohns(ylab = "Average Z (ages 5+)", col = pred_col)

m_plot <- trans_est(retro_tabs$log_aveM) %>%
  plot_mohns(ylab = "Average M (ages 5+)", col = pred_col)

f_plot <- trans_est(retro_tabs$log_aveF) %>%
  plot_mohns(ylab = "Average F (ages 5+)", col = pred_col)

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
  layout(annotations = left_label("Relative difference", offset = 50),
         margin = list(l = 75))

```

> Relative difference = (retro estimate - current estimate) / current estimate; Mohn's rho = mean(relative difference)

### Difference

```{r, eval = !is.null(fit$retro)}

z_plot <- retro_tabs$log_aveZ %>%
  plot_retro_deltas(ylab = "log(Average Z)", in_col = pred_col)

m_plot <- retro_tabs$log_aveM %>%
  plot_retro_deltas(ylab = "log(Average M)", in_col = pred_col)

f_plot <- retro_tabs$log_aveF %>%
  plot_retro_deltas(ylab = "log(Average F)", in_col = pred_col)

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
  layout(annotations = left_label("Difference", offset = 50),
         margin = list(l = 75))

```

> Difference = retro estimate - current estimate; approximate 95% intervals based on retro standard deviation + current standard deviation

Comps {data-navmenu="Trends" `r if (is.null(tmp$comp)) ".hidden"`}
================================================================================

Row
--------------------------------------------------------------------------------

### Stock status

```{r, eval = !is.null(tmp$comp)}

abn_plot <- trans_est(comp_tabs$log_recruitment, scale = 0.001) %>% 
  plot_comps(ylab = "Recruitment (age 2)", col = cols)

bms_plot <- trans_est(comp_tabs$log_ssb) %>% 
  plot_comps(ylab = "SSB (t)", col = cols, showlegend = FALSE)

blim_plot <- trans_est(comp_tabs$log_ssb_lrp) %>%
  plot_comps(ylab = "SSB/Blim", col = cols, showlegend = FALSE) %>%
  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none", showlegend = FALSE)

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

Row
--------------------------------------------------------------------------------

### Mortality Rates

```{r, eval = !is.null(tmp$comp)}

z_plot <- trans_est(comp_tabs$log_aveZ) %>%
  plot_comps(ylab = "Average Z (ages 5+)", col = cols)

m_plot <- trans_est(comp_tabs$log_aveM) %>%
  plot_comps(ylab = "Average M (ages 5+)", col = cols, showlegend = FALSE)

f_plot <- trans_est(comp_tabs$log_aveF) %>%
  plot_comps(ylab = "Average F (ages 5+)", col = cols, showlegend = FALSE)

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

Status and rates {data-navmenu="Trends"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Stock status

```{r}

## Note: biomass is in tonnes and numbers are in thousands
bms <- rbind(data.frame(metric = "SSB", tabs$log_ssb),
             data.frame(metric = "Ages 2+", tabs$log_biomass))
abn <- rbind(data.frame(metric = "Age 2", tabs$log_recruitment),
             data.frame(metric = "Ages 2+", tabs$log_abundance))

bms_plot <- trans_est(bms) %>%
  plot_trend(ylab = "Biomass (t)", group = "metric", col = cols,
             annotate = TRUE)

abn_plot <- trans_est(abn, scale = 0.001) %>%
  plot_trend(ylab = "Abundance", group = "metric", col = cols,
             annotate = TRUE)

blim_plot <- trans_est(tabs$log_ssb_lrp) %>%
  plot_trend(ylab = "SSB/Blim", annotate = TRUE, col = cols) %>%
  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none") %>%
  add_text(x = max_year + 1.2, y = 1, text = paste(round(fit$rep$blim / 1000), "kt"),
           color = I("black"), textposition = "right", hoverinfo = "none")

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

### Abundance at age

```{r}
plot_surface(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, name = "N")
```

<!-- ### Biomass at age -->

<!-- ```{r} -->
<!-- plot_surface(fit$rep$biomass_matrix, ages = ages, years = years_plus1, name = "Biomass") -->
<!-- ``` -->

### SSB at age

```{r}
plot_surface(fit$rep$ssb_matrix, ages = ages, years = years_plus1, name = "SSB")
```

<!-- ### Exploitable biomass at age -->

<!-- ```{r} -->
<!-- S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x)) -->
<!-- B <- fit$rep$N_matrix[, seq(ncol(S))] * fit$tmb.data$midyw -->
<!-- E <- B * S -->
<!-- plot_surface(E, ages = ages, years = years, name = "Exploitable biomass") -->
<!-- ``` -->

### Demographics

```{r}

a <- tidy_mat(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, value = "abundance") %>% 
  dplyr::left_join(trans_est(tabs$log_abundance, scale = 0.001), by = "year") %>% 
  rename(tot_abundance = est, tot_abundance_lwr = lwr, tot_abundance_upr = upr,
         tot_abundance_sd = sd)

b <- tidy_mat(fit$rep$biomass_matrix, ages = ages, years = years_plus1, value = "biomass") %>% 
  dplyr::left_join(trans_est(tabs$log_biomass), by = "year") %>% 
  rename(tot_biomass = est, tot_biomass_lwr = lwr, tot_biomass_upr = upr,
         tot_biomass_sd = sd)

c <- tidy_mat(fit$rep$ssb_matrix, ages = ages, years = years_plus1, value = "ssb") %>% 
  dplyr::left_join(trans_est(tabs$log_ssb), by = "year") %>% 
  dplyr::rename(tot_ssb = est, tot_ssb_lwr = lwr, tot_ssb_upr = upr,
                tot_ssb_sd = sd)

d <- dplyr::left_join(a, b, by = c("year", "age")) %>%
  dplyr::left_join(c, by = c("year", "age")) %>%
  mutate(cohort = year - age,
         abundance_prop = abundance / tot_abundance,
         abundance_lab = paste0("Abundance<br>", 
                                ifelse(gsub("[[:digit:]]", "", text_units(tot_abundance_lwr)) ==
                                         gsub("[[:digit:]]", "", text_units(tot_abundance_upr)),
                                       gsub("[[:alpha:]]", "", text_units(tot_abundance_lwr)),
                                       text_units(tot_abundance_lwr)), "-", 
                                text_units(tot_abundance_upr)),
         biomass_prop = -(biomass / tot_biomass),
         biomass_lab = paste0("Biomass<br>", 
                              ifelse(gsub("[[:digit:]]", "", text_units(tot_biomass_lwr)) ==
                                       gsub("[[:digit:]]", "", text_units(tot_biomass_upr)),
                                     gsub("[[:alpha:]]", "", text_units(tot_biomass_lwr)),
                                     paste0(text_units(tot_biomass_lwr), "t")), "-", 
                              text_units(tot_biomass_upr), "t"),
         ssb_prop = -(ssb / tot_ssb),
         ssb_lab = paste0("SSB<br>", 
                          ifelse(gsub("[[:digit:]]", "", text_units(tot_ssb_lwr)) ==
                                   gsub("[[:digit:]]", "", text_units(tot_ssb_upr)),
                                 gsub("[[:alpha:]]", "", text_units(tot_ssb_lwr)),
                                 paste0(text_units(tot_ssb_lwr), "t")), "-", 
                          text_units(tot_ssb_upr), "t")) %>%
  arrange(year, age)

max_x <- max(c(d$abundance_prop, d$biomass_prop)) * 1.2
xlabs <- pretty(c(-max_x, max_x), n = 10)

rec <- fit$rep$N_matrix[1, ]
rec_spikes <- find_peaks(rec, m = 1)
rec_spikes <- years[rec_spikes] - min(ages)

d$spike <- d$cohort
d$spike_start <- d$biomass_prop
d$spike_end <- d$abundance_prop
d$spike[!(d$spike %in% rec_spikes)] <- " "
d$spike_start[!(d$spike %in% rec_spikes)] <- 0
d$spike_end[!(d$spike %in% rec_spikes)] <- 0

d %>%
  plot_ly(y = ~age, frame = ~year, showlegend = FALSE,
          hoverinfo = "none", cliponaxis = FALSE) %>%
  add_paths(x = ~abundance_prop, fill = "tozerox",
            color = I(cols[1]), line = list(width = 1)) %>%
  # add_paths(x = ~biomass_prop, fill = "tozerox",
  #           color = I(cols[2]), line = list(width = 1)) %>%
  add_paths(x = ~ssb_prop, fill = "tozerox",
            color = I(cols[2]), line = list(width = 1)) %>%
  add_text(x = max_x, text = ~cohort, textposition = "right") %>%
  add_text(x = ~abundance_prop + 0.01, color = I("grey50"),
           text = ~spike, textposition = "center right") %>%
  add_text(x = ~ssb_prop - 0.01, color = I("grey50"),
           text = ~spike, textposition = "center left") %>%
  add_text(x = 0, y = ~max(age), text = ~paste0(year, "<br>"),
           textposition = "center top", size = I(18)) %>%
  # add_text(x = max_x / 2, y = ~max(age), text = "Abundance<br>",
  #          textposition = "center top", color = I(cols[1]),
  #          size = I(18)) %>%
  add_text(x = max_x / 2, y = ~max(age), text = ~abundance_lab,
           textposition = "center top", color = I(cols[1]),
           size = I(18)) %>%
  # add_text(x = -max_x / 2, y = ~max(age), text = ~biomass_lab,
  #          textposition = "center top", color = I(cols[2]),
  #          size = I(18)) %>%
  add_text(x = -max_x / 2, y = ~max(age), text = ~ssb_lab,
           textposition = "center top", color = I(cols[2]),
           size = I(18)) %>%
  add_annotations(x = 0, y = 1, yshift = 35, text = "Age",
                  xanchor = "right", yanchor = "top",
                  font = list(size = 14),
                  xref = "paper", yref = "paper", showarrow = FALSE) %>%
  add_annotations(x = 1, y = 1, yshift = 35, xshift = -5,
                  text = "Cohort",
                  xanchor = "left", yanchor = "top",
                  font = list(size = 14),
                  xref = "paper", yref = "paper", showarrow = FALSE) %>%
  layout(xaxis = list(range = c(-max_x, max_x),
                      title = "Percent",
                      tickmode = "array",
                      tickvals = xlabs,
                      ticktext = paste0(abs(xlabs) * 100, "%")),
         yaxis = list(title = ""),
         margin = list(l = 50, r = 70, t = 50, b = 50)) %>%
  animation_opts(frame = 1000, transition = 0) %>%
  animation_slider(currentvalue = list(prefix = "Year: ",
                                       font = list(color = "transparent")))

```

> Totals are lower and upper 95% confidence intervals

Row {.tabset}
--------------------------------------------------------------------------------

### Mortality Rates

```{r}

f <- data.frame(metric = "Ages 5+", tabs$log_aveF)
z <- data.frame(metric = "Ages 5+", tabs$log_aveZ)
m <- data.frame(metric = "Ages 5+", tabs$log_aveM)

z_plot <- trans_est(z) %>%
  plot_trend(ylab = "Average Z", group = "metric",
             annotate = TRUE, col = cols)

m_plot <- trans_est(m) %>%
  plot_trend(ylab = "Average M", group = "metric",
             annotate = TRUE, col = cols)

f_plot <- trans_est(f) %>%
  plot_trend(ylab = "Average F", group = "metric",
             annotate = TRUE, col = cols)

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

<!-- ### Z at age -->

<!-- ```{r} -->
<!-- plot_surface(fit$rep$Z_matrix, ages = ages, years = years, name = "Z") -->
<!-- ``` -->

### M at age

```{r}
plot_surface(fit$rep$Mpe_matrix, ages = ages, years = years, name = "M")
```

### F at age

```{r}
plot_surface(fit$rep$F_matrix, ages = ages, years = years, name = "F")
```

### Z, M and F

```{r}
plot_zmf_surface(z = fit$rep$Z_matrix, m = fit$rep$Mpe_matrix, f = fit$rep$F_matrix,
                 ages = ages, years = years)
```

> Z = purples, M = blues, F = reds

### Process error

```{r}

pe <- fit$rep$pe
plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], 
             ages = fit$tmb.data$pe_ages, name = "pe", 
             col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
  add_surface(matrix(0, nrow = nrow(pe), ncol = ncol(pe)), showscale = FALSE,
              color = I("black"), opacity = 0.8)

```

<!-- ### Process error (bubble) -->

<!-- ```{r} -->
<!-- pe <- tidy_mat(fit$rep$pe, years = c(years, fit$tmb.data$proj_years)[-1], -->
<!--                ages = fit$tmb.data$pe_ages, value = "res") -->
<!-- bubble_plot(pe, pos_col = pos_col, neg_col = neg_col) -->
<!-- ``` -->

Productivity  {data-navmenu="Trends"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Recruitment

```{r}
trans_est(tabs$log_recruitment, scale = 0.001) %>%
  plot_trend(ylab = "Recruitment", col = pred_col)
```

### SSB

```{r}
trans_est(tabs$log_ssb) %>%
  plot_trend(ylab = "SSB (t)", col = pred_col)
```

### RPS

```{r}
trans_est(tabs$log_rps) %>%
  filter(year >= min_year + min(ages)) %>% 
  plot_trend(ylab = "RPS (thousand / tonne)", col = pred_col) %>% 
  layout(xaxis = list(range = c(min_year, max_year + 1)))
```

### SSB x Recruitment

```{r}

ssb <- trans_est(tabs$log_ssb)
rec <- trans_est(tabs$log_recruitment, scale = 0.001)
rps <- trans_est(tabs$log_rps)
names(ssb) <- paste0("ssb_", names(ssb))
names(rec) <- paste0("rec_", names(rec))
names(rps) <- paste0("rps_", names(rps))
ssb$year <- ssb$ssb_year + min(ages)
rec$year <- rec$rec_year
rps$year <- rps$rps_year
sr <- merge(ssb, rec, by = "year")
sr <- merge(sr, rps, by = "year")
sr$lab <- paste0("Recruitment year: ", sr$rec_year, "<br>SSB year: ", sr$ssb_year)

plot_xy_errorbar(sr, x = ~ssb_est, y = ~rec_est,
                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
                 y_lwr = ~rec_lwr, y_upr = ~rec_upr,
                 text = ~lab, xlab = "SSB (t)", ylab = "Recruitment",
                 col = ~year, clab = "Year")

```

### SSB x RPS

```{r}
p <- plot_xy_errorbar(sr, x = ~ssb_est, y = ~rps_est,
                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
                 y_lwr = ~rps_lwr, y_upr = ~rps_upr,
                 text = ~lab, xlab = "SSB (t)", ylab = "RPS (thousand / tonne)",
                 col = ~year, clab = "Year")
p
```

### log(SSB) x log(RPS)

```{r}
p %>% layout(yaxis = list(type = "log"),
             xaxis = list(type = "log"))
```